\chapter{Use-Case Analysis}

This chapter describes the actors that interact with the Mini Payment Gateway
and the main use cases the system supports. The focus is the business role of
each actor and the expected system outcome.

\section{Actor Overview}

\begin{table}[H]
\centering
\caption{Implemented actors}
\begin{tabularx}{\textwidth}{L{0.25\textwidth}Y}
\toprule
\textbf{Actor} & \textbf{Role in the system} \\
\midrule
Admin & Privileged internal user who manages internal accounts, merchant
readiness, merchant credentials, and merchant dashboard accounts. \\
Ops & Internal operations user who monitors transactions, investigates delivery
issues, reviews evidence, and performs allowed recovery actions. \\
Merchant backend & Merchant-owned application that creates payments and refunds
through the gateway on behalf of the merchant business. \\
Merchant dashboard user & Human merchant representative who signs in to inspect
transactions, analytics, profile information, and credential metadata. \\
Provider callback source & External payment-result source that confirms payment
and refund outcomes. \\
Scheduled trigger & Time-based trigger that activates expiration handling and
webhook retry processing. \\
\bottomrule
\end{tabularx}
\end{table}

\section{Core Gateway Use Cases}

\begin{longtable}{L{0.11\textwidth}L{0.25\textwidth}L{0.18\textwidth}L{0.34\textwidth}}
\caption{Core gateway use cases}\\
\toprule
\textbf{ID} & \textbf{Use case} & \textbf{Primary actor} & \textbf{System outcome} \\
\midrule
\endfirsthead
\toprule
\textbf{ID} & \textbf{Use case} & \textbf{Primary actor} & \textbf{System outcome} \\
\midrule
\endhead
UC001 & Onboard and activate merchant & Admin or Ops & Merchant profile, review evidence, access credentials, and operational status are established before live transaction activity. \\
UC002 & Create dynamic QR payment & Merchant backend & A pending payment request is created and a payment payload is returned for customer-facing checkout. \\
UC003 & Process payment result callback & Provider callback source & Callback evidence is stored and the payment moves to a final state or to manual review when evidence is ambiguous. \\
UC004 & Create full refund and process result & Merchant backend and provider callback source & A full refund request is created and later finalized from external confirmation. \\
UC005 & Deliver webhook with retry and manual recovery & Scheduled trigger and internal operator & Merchant notification attempts are recorded, retries are controlled, and failed deliveries can be retried safely. \\
\bottomrule
\end{longtable}

The following diagram summarizes the core gateway use cases only. Dashboard
screens are intentionally excluded so the diagram stays focused on the main
money-flow and notification responsibilities of the gateway itself.

\begin{figure}[H]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[node distance=11mm and 12mm]
  \node[actor] (admin) {Admin / Ops};
  \node[actor, below=of admin] (merchant) {Merchant backend};
  \node[actor, below=of merchant] (provider) {Provider callback\\source};
  \node[actor, below=of provider] (trigger) {Scheduled\\trigger};

  \node[boundary, right=28mm of merchant, minimum width=45mm, minimum height=31mm, label=above:{Merchant readiness}] (readinessBox) {};
  \node[boundary, right=of readinessBox, minimum width=55mm, minimum height=68mm, label=above:{Money movement lifecycle}] (moneyBox) {};
  \node[boundary, below right=4mm and 6mm of moneyBox, minimum width=58mm, minimum height=39mm, label=above:{Notification and evidence}] (notifyBox) {};

  \node[box] at ([yshift=2mm]readinessBox.center) (onboard) {Onboard and activate merchant};
  \node[box] at ([yshift=16mm]moneyBox.center) (pay) {Create dynamic QR payment};
  \node[box] at ([yshift=-1mm]moneyBox.center) (callback) {Process payment result callback};
  \node[box] at ([yshift=-18mm]moneyBox.center) (refund) {Create full refund and\\process refund result};
  \node[box] at ([yshift=8mm]notifyBox.center) (webhook) {Deliver webhook with retry\\and manual recovery};
  \node[box] at ([yshift=-13mm]notifyBox.center) (review) {Review audit and\\reconciliation evidence};

  \draw[arrow] (admin.east) -- (onboard.west);
  \draw[arrow] (admin.east) to[out=8,in=170] (review.west);
  \draw[arrow] (merchant.east) -- (pay.west);
  \draw[arrow] (merchant.east) to[out=6,in=182] (refund.west);
  \draw[arrow] (provider.east) -- (callback.west);
  \draw[arrow] (provider.east) to[out=3,in=192] (refund.west);
  \draw[arrow] (trigger.east) to[out=5,in=205] (webhook.west);
  \draw[arrow] (callback.east) -- ++(11mm,0) |- (webhook.west);
  \draw[arrow] (refund.east) -- ++(11mm,0) |- (webhook.west);
\end{tikzpicture}
}
\caption{Core gateway use-case diagram. The diagram focuses on merchant
readiness, payment or refund state changes, and notification or evidence
handling inside the gateway.}
\end{figure}

\section{Dashboard Use Cases}

\begin{longtable}{L{0.11\textwidth}L{0.26\textwidth}L{0.18\textwidth}L{0.27\textwidth}}
\caption{Dashboard use cases}\\
\toprule
\textbf{ID} & \textbf{Use case} & \textbf{Actor} & \textbf{System outcome} \\
\midrule
\endfirsthead
\toprule
\textbf{ID} & \textbf{Use case} & \textbf{Actor} & \textbf{System outcome} \\
\midrule
\endhead
DUC001 & Sign in and view operations overview & Admin or Ops & Internal users access a consolidated operational view of merchants, transactions, delivery issues, and pending reviews. \\
DUC002 & Manage merchant lifecycle & Admin or Ops & Internal users register merchants, review readiness information, manage credentials, and apply lifecycle decisions. \\
DUC003 & Manage merchant dashboard accounts & Admin & Merchant-facing user accounts can be created, updated, deactivated, reactivated, or reset. \\
DUC004 & Inspect operational evidence & Admin or Ops & Internal users review payments, refunds, webhook history, reconciliation items, and audit evidence. \\
DUC005 & Sign in and maintain account access & Merchant dashboard user & Merchant representatives authenticate securely and can change their own password. \\
DUC006 & View overview and analytics & Merchant dashboard user & Merchant representatives view summarized activity and trend information for a selected reporting window. \\
DUC007 & Explore payment, refund, and webhook records & Merchant dashboard user & Merchant representatives review transaction and notification history that belongs only to their merchant account. \\
DUC008 & View profile and credential metadata & Merchant dashboard user & Merchant representatives can read merchant profile information and masked credential details without seeing raw secrets. \\
\bottomrule
\end{longtable}
