\chapter{Database Design}

\section{Entity Model}

The data model is designed around durable business evidence. It must support
merchant readiness, transaction state, refund state, notification delivery,
external outcome evidence, reconciliation, audit history, and human account
access without losing traceability.

These entities fall into four practical groups: merchant identity and access,
money-flow records, external delivery evidence, and operational traceability.
The overview below emphasizes those relationships in a portrait-friendly form so
the reader can connect the major business objects without needing a landscape
schema page.

\begin{figure}[!htbp]
\centering
\resizebox{0.92\textwidth}{!}{%
\begin{tikzpicture}[node distance=9mm and 12mm]
  \node[box, text width=31mm] (internalUser) {Internal user};
  \node[box, right=of internalUser, text width=31mm] (merchant) {Merchant};
  \node[box, right=of merchant, text width=33mm] (merchantUser) {Merchant dashboard user};

  \node[box, below=of internalUser, text width=31mm] (onboarding) {Onboarding case};
  \node[box, below=of merchant, text width=31mm] (credential) {Merchant credential};
  \node[box, below=of merchantUser, text width=33mm] (payment) {Payment transaction};

  \node[box, below=of onboarding, text width=31mm] (audit) {Audit log};
  \node[box, below=of credential, text width=31mm] (callback) {Callback evidence};
  \node[box, below=of payment, text width=33mm] (webhook) {Webhook event};

  \node[box, below left=of webhook, text width=31mm] (refund) {Refund transaction};
  \node[box, below=of webhook, text width=33mm] (attempt) {Webhook delivery attempt};
  \node[box, below left=of attempt, text width=31mm] (recon) {Reconciliation record};

  \draw[arrow] (internalUser) -- (onboarding);
  \draw[arrow] (internalUser) -- (audit);
  \draw[arrow] (merchant) -- (onboarding);
  \draw[arrow] (merchant) -- (credential);
  \draw[arrow] (merchant) -- (merchantUser);
  \draw[arrow] (merchant) -- (payment);
  \draw[arrow] (merchant) -- (audit);
  \draw[arrow] (merchantUser) -- (audit);
  \draw[arrow] (payment) -- (refund);
  \draw[arrow] (payment) -- (webhook);
  \draw[arrow] (webhook) -- (attempt);
  \draw[arrow] (callback) -- (payment);
  \draw[arrow] (callback) -- (refund);
  \draw[arrow] (callback) -- (recon);
  \draw[arrow] (payment) -- (recon);
  \draw[arrow] (refund) -- (recon);
\end{tikzpicture}
}
\caption{Logical entity relationship overview. The model centers on merchants,
transactions, delivery evidence, and operational traceability.}
\end{figure}

\section{Core Entities}

\begin{longtable}{L{0.26\textwidth}L{0.64\textwidth}}
\caption{Core business entities}\\
\toprule
\textbf{Entity} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Entity} & \textbf{Purpose} \\
\midrule
\endhead
Merchant & Stores the merchant identity, operating profile, contact data, notification destination, settlement information, and current operating status. \\
\bodyrowrule
Onboarding Case & Stores merchant-submitted readiness information, review findings, approval or rejection outcomes, and review history. \\
\bodyrowrule
Merchant Credential & Stores merchant integration credential metadata and protected secret material used for signed transaction requests. \\
\bodyrowrule
Merchant Dashboard User & Stores human merchant accounts that sign in to the merchant dashboard and are scoped to one merchant. \\
\bodyrowrule
Internal User & Stores internal administrative and operational accounts used for management and review workflows. \\
\bodyrowrule
Payment Transaction & Stores payment attempts, money amounts, state, expiration information, merchant references, and externally linked payment evidence. \\
\bodyrowrule
Refund Transaction & Stores refund requests and their linkage to successful original payments. \\
\bodyrowrule
Webhook Event & Stores merchant notification payloads, delivery state, retry metadata, and notification history. \\
\bodyrowrule
Webhook Delivery Attempt & Stores individual delivery attempts so the system can preserve delivery evidence over time. \\
\bodyrowrule
Callback Evidence & Stores raw and normalized external outcome evidence for payments and refunds. \\
\bodyrowrule
Reconciliation Record & Stores manual-review evidence when external outcomes do not align cleanly with internal state. \\
\bodyrowrule
Audit Log & Stores durable review and administrative evidence for sensitive human actions. \\
\bottomrule
\end{longtable}

\section{Data Invariants}

\begin{longtable}{L{0.34\textwidth}L{0.56\textwidth}}
\caption{Important data invariants}\\
\toprule
\textbf{Invariant} & \textbf{Reason} \\
\midrule
\endfirsthead
\toprule
\textbf{Invariant} & \textbf{Reason} \\
\midrule
\endhead
Each merchant has one stable public merchant identity. & External integrations and both dashboards need one durable merchant anchor. \\
\bodyrowrule
Only one active merchant integration credential exists at a time. & Signed requests should resolve to one authoritative active credential for each merchant. \\
\bodyrowrule
Each merchant dashboard user belongs to exactly one merchant. & Merchant dashboard access must remain tenant-scoped and must not leak across merchants. \\
\bodyrowrule
Only one active pending payment can exist for the same merchant order at a time. & This prevents duplicate active checkout attempts for one business order. \\
\bodyrowrule
Refund requests are unique per merchant refund reference. & The system needs business-level idempotency for merchant refund initiation. \\
\bodyrowrule
At most one successful full refund can exist for one successful payment. & The project implements full-refund-only behavior. \\
\bodyrowrule
Notification delivery history is preserved separately from notification state. & Final notification status and the evidence that led to it must both remain available. \\
\bodyrowrule
Protected secrets and plaintext passwords are never returned through ordinary read views. & Human access and merchant integration credentials must remain protected even when metadata is viewable. \\
\bottomrule
\end{longtable}

\section{Data Evolution}

The data model evolves by extending business capabilities rather than by
rewriting existing evidence. As the system grows, new entities or attributes
can be introduced for additional merchant lifecycle information, expanded
notification channels, or richer reconciliation support while preserving
historical transaction and audit integrity.
