\chapter{API Design and Implementation}

\section{API Documentation Strategy}

The main report uses compact API family IDs so that the design discussion stays
readable. The companion document in \code{api-documentation.tex} owns the
detailed
request and response contracts, shared conventions, error cases, security
notes, and Postman scenario mapping.

This chapter therefore focuses on API design decisions rather than endpoint-by-
endpoint payload detail.

\section{API Taxonomy And Boundary Split}

\begin{table}[H]
\centering
\caption{API families and access boundaries}
\begin{tabularx}{\textwidth}{L{0.18\textwidth}L{0.20\textwidth}Y}
\toprule
\textbf{Family} & \textbf{Actor or auth} & \textbf{Design purpose} \\
\midrule
\apiid{AUTH-01} & Internal Admin/Ops session & Bootstrap, login, logout, current-session lookup, password change, and internal user administration support. \\
\apiid{AUTH-02--03} & Merchant portal session & Merchant Dashboard login, logout, current-session lookup, and password change. \\
\apiid{PAY-*} and \apiid{REF-*} & Merchant backend / HMAC & Merchant-facing payment and refund creation plus transaction lookup. \\
\apiid{CB-*} and \apiid{REF-04--05} & Provider simulator & Provider-side payment and refund result callbacks with reconciliation fallback. \\
\apiid{OPS-*}, \apiid{ONB-*}, \apiid{MPU-*}, \apiid{REC-*} & Internal Admin/Ops session & Internal dashboard reads, merchant lifecycle actions, portal-user provisioning, reconciliation, and recovery. \\
\apiid{MP-*} & Merchant portal session & Merchant-scoped overview, analytics, explorers, profile, and credentials. \\
\apiid{WH-*} & Gateway-to-merchant plus internal Ops retry & Outbound webhook event contract, retry behavior, and manual recovery. \\
\bottomrule
\end{tabularx}
\end{table}

The implemented API design uses three separate authentication models rather than
one shared credential system:

\begin{itemize}
  \item merchant backends call \apiid{PAY-*} and \apiid{REF-*} with HMAC
  headers;
  \item internal users call \apiid{AUTH-01}, \apiid{OPS-*}, \apiid{ONB-*},
  \apiid{MPU-*}, and \apiid{REC-*} with an internal HttpOnly session and RBAC;
  \item Merchant Dashboard users call \apiid{AUTH-02--03} and \apiid{MP-*}
  with a merchant portal HttpOnly session.
\end{itemize}

\section{Reduced API ID Catalogue}

\begin{longtable}{L{0.16\textwidth}L{0.27\textwidth}L{0.18\textwidth}L{0.29\textwidth}}
\caption{Reduced API catalogue for the report narrative}\\
\toprule
\textbf{ID family} & \textbf{Representative routes} & \textbf{Actor/auth} & \textbf{Design note} \\
\midrule
\endfirsthead
\toprule
\textbf{ID family} & \textbf{Representative routes} & \textbf{Actor/auth} & \textbf{Design note} \\
\midrule
\endhead
\apiid{AUTH-01} & \routecell{\texttt{/v1/internal/auth/*}\\\texttt{/v1/internal/users/*}} & Admin/Ops session & Internal auth is session-based and supports RBAC plus internal user administration. \\
\apiid{PAY-01--03} & \routecell{\texttt{POST /v1/payments}\\\texttt{GET /v1/payments/*}} & Merchant backend / HMAC & Payment creation enforces merchant readiness and semantic idempotency. \\
\apiid{REF-01--05} & \routecell{\texttt{POST /v1/refunds}\\\texttt{GET /v1/refunds/*}\\\texttt{/v1/provider/callbacks/refund}} & Merchant backend or provider simulator & Refund design is full-refund-only and integrates callback-based finalization. \\
\apiid{CB-01--02} & \routecell{\texttt{/v1/provider/callbacks/payment}} & Provider simulator & Ambiguous or mismatched evidence produces reconciliation rather than silent overwrite. \\
\apiid{OPS-01--06} & \routecell{\texttt{/v1/ops/dashboard/*}\\\texttt{/v1/ops/merchants/*}} & Admin/Ops session & Internal dashboards combine read APIs, merchant lifecycle actions, credentials, and internal user workflows. \\
\apiid{MPU-01--04} & \routecell{\texttt{/v1/ops/merchants/}\\\texttt{.../portal-users/*}} & Admin session & Merchant portal user provisioning is internal-only and returns generated passwords once. \\
\apiid{MP-01--06} & \routecell{\texttt{/v1/merchant-portal/*}} & Merchant portal session & Merchant APIs are read-only, merchant-scoped, and never trust client-supplied merchant ids after login. \\
\apiid{REC-01--02} and \apiid{WH-10} & \routecell{\texttt{/v1/ops/reconciliation/*}\\\texttt{/v1/ops/webhooks/*/retry}} & Admin/Ops session & Recovery APIs support review and manual retry without mutating already-final money-flow state. \\
\bottomrule
\end{longtable}

\section{Asynchronous API And Event Design}

The API design must handle asynchronous evidence safely. Provider callbacks do
not merely set a status field; they first store durable evidence and then apply
domain rules. Late success after expiration, amount mismatch, and final-state
conflicts create reconciliation records instead of forcing incorrect state
transitions.

Webhook delivery is intentionally decoupled from payment and refund finality.
The API surface exposes the business event indirectly through webhook event
creation and directly through internal retry support. This makes it possible to
recover from merchant endpoint failure without changing the final payment or
refund outcome.

\section{Error Model And Security Design}

Business failures use stable application error codes rather than ad hoc status
messages. Examples include \texttt{MERCHANT\_NOT\_ACTIVE},
\texttt{ACTIVE\_CREDENTIAL\_REQUIRED}, \texttt{PAYMENT\_ALREADY\_SUCCESS}, and
\texttt{MERCHANT\_AUTH\_INACTIVE}.

Sensitive data exposure is minimized:

\begin{itemize}
  \item raw credential secrets are never returned by credential metadata APIs;
  \item generated portal passwords are returned only on create and reset;
  \item audit snapshots mask secret-like fields recursively;
  \item Merchant Portal detail routes return 404 for records outside the
  authenticated merchant scope.
\end{itemize}

The detailed payload examples, common error envelope, and Postman mapping are
maintained in the companion API documentation so the report can stay design-
focused.
