\documentclass[11pt,a4paper,oneside]{report}

\newcommand{\DocumentTitle}{System Design and Implementation Report}
\newcommand{\DocumentSummary}{This report documents the implemented Mini Payment Gateway as an academic project artifact. It covers requirements, architecture, database design, user interface design, implementation structure, API design, and testing evidence while staying aligned with the behavior currently present in the repository.}

\input{shared/metadata.tex}
\input{shared/preamble.tex}
\input{shared/macros.tex}

\begin{document}

\hypersetup{pageanchor=false}
\makempgtitlepage
\clearpage
\hypersetup{pageanchor=true}

\pagenumbering{roman}

\frontchapter{Abstract}
The Mini Payment Gateway is an implemented MVP payment gateway for merchant
onboarding, dynamic QR payment creation, provider result callbacks, full
refunds, webhook delivery, operational investigation, and dashboard-based
merchant visibility. The system combines a FastAPI backend, PostgreSQL
database, an internal Ops Dashboard, and a separate Merchant Dashboard.

The report is intentionally implementation-faithful. It focuses on the current
repository state rather than speculative features, and it uses the companion
API document in \code{api-documentation.tex} for the detailed endpoint
contracts.

\tableofcontents
\listoffigures
\listoftables

\frontchapter{Glossary and Acronyms}
\begin{longtable}{L{0.20\textwidth}L{0.70\textwidth}}
\toprule
\textbf{Term} & \textbf{Meaning} \\
\midrule
\endfirsthead
\toprule
\textbf{Term} & \textbf{Meaning} \\
\midrule
\endhead
API & Application Programming Interface. \\
E2E & End-to-end verification across multiple components. \\
HMAC & Hash-based message authentication code used by merchant-facing APIs. \\
MVP & Minimum viable product scope implemented for the project. \\
Ops & Internal operations role used for day-to-day support workflows. \\
RBAC & Role-based access control for internal and merchant-facing session users. \\
QR & Quick response payment payload returned to merchants during payment creation. \\
\bottomrule
\end{longtable}

\clearpage
\pagenumbering{arabic}

\input{chapters/01-system-requirements-and-problem-analysis.tex}
\input{chapters/02-use-case-analysis.tex}
\input{chapters/03-user-stories.tex}
\input{chapters/04-system-architecture.tex}
\input{chapters/05-component-architecture-and-design.tex}
\input{chapters/06-database-design.tex}
\input{chapters/07-user-interface-design.tex}
\input{chapters/08-component-implementation.tex}
\input{chapters/09-api-design-and-implementation.tex}
\input{chapters/10-testing-and-evaluation.tex}

\appendix
\input{appendices/demonstration-storyline.tex}
\input{appendices/limitations-and-future-work.tex}

\end{document}
