\chapter{User Stories}

The user stories are grouped by actor so that the report can connect user
intent to the implemented workflows and boundaries.

\begin{longtable}{L{0.17\textwidth}L{0.37\textwidth}L{0.36\textwidth}}
\caption{Implemented user stories}\\
\toprule
\textbf{ID} & \textbf{Story} & \textbf{Acceptance criteria} \\
\midrule
\endfirsthead
\toprule
\textbf{ID} & \textbf{Story} & \textbf{Acceptance criteria} \\
\midrule
\endhead
US-ADMIN-01 & As an Admin, I want to create and review merchants so that only approved merchants can use the gateway. & Admin can create a merchant, submit and review onboarding evidence, create credentials, and activate the merchant. \\
\bodyrowrule
US-ADMIN-02 & As an Admin, I want to manage internal users so that operations access can be controlled. & Admin can list, create, update, deactivate or reactivate, and reset internal user passwords. \\
\bodyrowrule
US-ADMIN-03 & As an Admin, I want to provision merchant dashboard users so that merchants can inspect their own gateway data. & Admin can create, update, deactivate or reactivate, and reset merchant portal users from merchant detail. \\
\bodyrowrule
US-OPS-01 & As an Ops user, I want to monitor operational queues so that failed or pending work can be handled quickly. & Ops can log in, view metrics and charts, and navigate to payments, refunds, webhooks, and reconciliation. \\
\bodyrowrule
US-OPS-02 & As an Ops user, I want to manage merchant lifecycle states safely. & Ops can perform allowed lifecycle actions with a required reason and audit trail. \\
\bodyrowrule
US-OPS-03 & As an Ops user, I want to inspect webhook attempts so failed notifications can be retried or explained. & Ops can open webhook detail, see attempts, and retry failed webhook events where allowed. \\
\bodyrowrule
US-OPS-04 & As an Ops user, I want to configure a merchant's VietQR receiving account so that payment QR generation is controlled and auditable. & Ops can list, create, activate, and deactivate merchant QR accounts from merchant detail, with one active account per provider and audit evidence for changes. \\
\bodyrowrule
US-OPS-05 & As an Ops user, I want to provision merchant dashboard users so merchant access can be completed during onboarding and support. & Ops can create, update, deactivate or reactivate, and reset merchant portal users; generated passwords are shown once and every mutation records the authenticated Ops actor. \\
\bodyrowrule
US-INTEGRATION-01 & As a merchant backend, I want payment creation to return a scannable QR image and stable QR reference. & A valid signed payment request for an active, QR-configured merchant returns \code{qr_reference}, \code{qr_content}, and \code{qr_image_base64}; duplicate equivalent pending requests return the same QR evidence. \\
\bodyrowrule
US-MERCHANT-01 & As a merchant operator, I want to log in without API credentials. & Merchant user logs in with merchant id, email, and password; the session is separate from HMAC API credentials. \\
\bodyrowrule
US-MERCHANT-02 & As a merchant operator, I want to see overview metrics. & Overview shows summary cards and compact charts for the logged-in merchant only. \\
\bodyrowrule
US-MERCHANT-03 & As a merchant operator, I want interactive analytics with exact values. & Analytics supports 7, 30, and 90 day ranges, tooltips, data tables, and drill-down links. \\
\bodyrowrule
US-MERCHANT-04 & As a merchant operator, I want to inspect payments, refunds, and webhooks for support questions. & Explorer pages list and show detail for only the logged-in merchant's records, including QR image and payload details when the payment has generated VietQR data. \\
\bodyrowrule
US-VIEWER-01 & As a merchant business viewer, I want to view revenue and success-rate trends. & Viewer can open Overview and Analytics, read chart values, and switch date ranges. \\
\bodyrowrule
US-VIEWER-02 & As a merchant business viewer, I want read-only credential metadata. & Credentials page shows access key, secret suffix, status, and timestamps only. \\
\bottomrule
\end{longtable}
