\section{Quickstart}

This guide is written for teams that need to integrate with the Mini Payment
Gateway without reading internal project materials. Public integration surfaces
appear first. Restricted operational endpoints appear later so merchant teams
can recognize them during onboarding, support, and recovery conversations.

\docbox{Recommended integration order}{%
\begin{enumerate}[leftmargin=1.7em]
  \item Receive your merchant identifier, active access key, and shared secret.
  \item Implement HMAC request signing, then create and query payments.
  \item Implement full-refund creation and refund status lookups.
  \item Prepare to receive gateway webhooks for final payment and refund events.
  \item If your environment simulates provider callbacks, connect the callback
  source in the controlled environment agreed during onboarding.
  \item Use the Merchant Dashboard only for human review, not for server-to-server
  payment processing.
\end{enumerate}
}

\begin{longtable}{@{}L{0.24\textwidth}L{0.68\textwidth}@{}}
\caption{Audience legend}\\
\toprule
\textbf{Audience} & \textbf{How this guide uses the term} \\
\midrule
\endfirsthead
\toprule
\textbf{Audience} & \textbf{How this guide uses the term} \\
\midrule
\endhead
Merchant backend & The merchant-owned server or application that signs payment
and refund API requests. \\
\bodyrowrule
Merchant dashboard user & A signed-in business user who reviews transactions,
refunds, webhooks, and analytics in the Merchant Dashboard. \\
\bodyrowrule
Provider callback source & A bank, provider, or simulator that sends payment or
refund result notifications into the gateway. \\
\bodyrowrule
Restricted / Internal operator & An internal support or operations user who
handles onboarding, account provisioning, reconciliation, and recovery. \\
\bottomrule
\end{longtable}

\apiblockheading{Implemented scope.}
This guide covers payment creation, payment lookup, full refunds, provider
result callbacks, outbound webhook delivery, Merchant Dashboard access, and the
restricted operational endpoints that support onboarding and support recovery.

\apiblockheading{Out of scope in the current MVP.}
The current product does not document settlement, dispute workflows, partial
refunds, merchant-run onboarding, or real-time push streams beyond the
documented webhook model.
