\section{Reference IDs}

Reference IDs such as \apiid{PAY-01} or \apiid{WH-10} are stable support labels.
They help merchants, implementation teams, and operators talk about the same
API family without repeating long endpoint lists in every discussion.

\begin{longtable}{@{}L{0.14\textwidth}L{0.26\textwidth}L{0.52\textwidth}@{}}
\caption{Reference ID prefixes}\\
\toprule
\textbf{Prefix} & \textbf{API family} & \textbf{Typical use} \\
\midrule
\endfirsthead
\toprule
\textbf{Prefix} & \textbf{API family} & \textbf{Typical use} \\
\midrule
\endhead
\apiid{AUTH} & Session authentication & Sign-in, sign-out, current-session,
and password-change flows for interactive users. \\
\bodyrowrule
\apiid{PAY} & Merchant payment APIs & Payment creation and payment status
lookup from the merchant backend. \\
\bodyrowrule
\apiid{REF} & Refund APIs & Refund creation, refund status lookup, and refund
result callback flows. \\
\bodyrowrule
\apiid{CB} & Provider callback APIs & Payment result callbacks from a bank,
provider, or simulator. \\
\bodyrowrule
\apiid{WH} & Webhook delivery & Outbound event delivery to the merchant's
webhook endpoint and operational retry actions. \\
\bodyrowrule
\apiid{OPS} & Restricted operations APIs & Internal dashboard reads, merchant
lifecycle actions, credential management, and support operations. \\
\bodyrowrule
\apiid{ONB} & Onboarding APIs & Submission and review of merchant onboarding
cases. \\
\bodyrowrule
\apiid{MPU} & Merchant dashboard user provisioning & Creation and maintenance
of Merchant Dashboard user accounts by internal operators. \\
\bodyrowrule
\apiid{MP} & Merchant dashboard APIs & Dashboard summary, analytics,
explorers, profile, and credential metadata. \\
\bodyrowrule
\apiid{REC} & Reconciliation APIs & Review and resolution of mismatches
between internal records and external evidence. \\
\bottomrule
\end{longtable}

These IDs are intended as reader-friendly labels. They do not change request
paths, response payloads, or authentication rules.
