\section{Authentication And Access Models}

\begin{longtable}{@{}L{0.21\textwidth}L{0.23\textwidth}L{0.22\textwidth}L{0.26\textwidth}@{}}
\caption{Authentication and scope matrix}\\
\toprule
\textbf{Audience} & \textbf{Primary surface} & \textbf{Authentication} &
\textbf{How scope is determined} \\
\midrule
\endfirsthead
\toprule
\textbf{Audience} & \textbf{Primary surface} & \textbf{Authentication} &
\textbf{How scope is determined} \\
\midrule
\endhead
Merchant backend & Merchant payment and refund APIs & HMAC headers carrying the
merchant identifier, access key, timestamp, and signature. & The active
merchant credential used to sign the request. \\
\bodyrowrule
Merchant dashboard user & Merchant Dashboard APIs & Signed-in dashboard session
cookie. & The merchant account linked to the signed-in dashboard user. \\
\bodyrowrule
Provider callback source & Provider callback APIs & Controlled-environment
trust in the current MVP. & The payment or refund reference carried in the
callback payload. \\
\bodyrowrule
Restricted / Internal operator & Internal and operations APIs & Signed-in
internal session cookie with role-based access rules. & The operator role and
identity carried by the authenticated internal session. \\
\bottomrule
\end{longtable}

\docbox{Boundary rules that matter most}{%
\begin{itemize}
  \item Merchant backend APIs and Merchant Dashboard APIs are separate surfaces.
  A dashboard user cannot replace HMAC credentials, and HMAC credentials cannot
  sign into the dashboard.
  \item Merchant Dashboard routes derive merchant scope from the signed-in
  account. Clients do not choose the merchant scope by sending a separate
  merchant identifier after login.
  \item Restricted internal write requests may include a human reason for audit
  purposes, but operator identity and permissions still come from the signed-in
  internal session.
  \item Provider callback routes are intended for controlled private
  environments in the current MVP and should not be treated as a public,
  internet-grade inbound provider integration without additional hardening.
\end{itemize}
}
