\section{Merchant Backend APIs}

Merchant backend APIs are machine-to-machine endpoints. They are not used by
human dashboard users. Every request in this section is signed with the
merchant's HMAC credential set.

\subsection{Payments}

\begin{longtable}{@{}L{0.25\textwidth}L{0.65\textwidth}@{}}
\caption{Payment API family summary}\\
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endfirsthead
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endhead
Purpose & Create payment intents and retrieve payment status for a merchant's
own orders. \\
\bodyrowrule
Intended audience & Merchant backend. \\
\bodyrowrule
Authentication & Merchant HMAC headers. \\
\bodyrowrule
Reference IDs & \apiid{PAY-01}, \apiid{PAY-02}, \apiid{PAY-03}. \\
\bottomrule
\end{longtable}

\apiblockheading{Endpoints.}
\begin{longtable}{@{}L{0.31\textwidth}L{0.59\textwidth}@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endhead
\code{POST /v1/payments} & Create a new pending payment and return the gateway
payment reference plus QR payload information. \\
\bodyrowrule
\code{GET /v1/payments/{transaction_id}} & Retrieve a payment by gateway
transaction identifier within the authenticated merchant scope. \\
\bodyrowrule
\code{GET /v1/payments/by-order/{order_id}} & Retrieve the latest payment
attempt for the merchant's order identifier. \\
\bottomrule
\end{longtable}

\apiblockheading{Important request fields.}
The payment creation request requires \code{order_id}, \code{amount}, and
\code{description}. The caller must send exactly one expiration strategy:
\code{ttl_seconds} or \code{expire_at}. Optional business metadata may be
attached in \code{metadata}.

\apiblockheading{Important response fields.}
The creation response highlights \code{transaction_id}, \code{order_id},
\code{merchant_id}, \code{qr_content}, optional QR image fields, the current
\code{status}, and \code{expire_at}.

\apiblockheading{Business rules and usage notes.}
\begin{itemize}
  \item The merchant account must be active before a payment can be created.
  \item The gateway stores a usable payment directly as \code{PENDING}; there is
  no separate client-visible pre-pending state.
  \item Only one active pending payment may exist for the same merchant order
  at a time.
  \item A semantically identical retry for the current pending order returns
  the existing payment instead of creating a new one.
  \item A different retry while a pending payment already exists is rejected.
  \item A new payment is allowed after a previous attempt has failed or
  expired.
  \item A new payment is rejected after the order already has a successful
  payment.
\end{itemize}

\apiblockheading{Likely error codes.}
\begin{longtable}{@{}L{0.28\textwidth}L{0.62\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endhead
\code{MERCHANT_NOT_ACTIVE} & The merchant is not currently allowed to create
new payment or refund activity. \\
\bodyrowrule
\code{PAYMENT_PENDING_EXISTS} & The same order already has a different pending
payment request that has not been resolved yet. \\
\bodyrowrule
\code{PAYMENT_ALREADY_SUCCESS} & The order has already been paid successfully
and should not be opened again as a new payment. \\
\bodyrowrule
\code{PAYMENT_NOT_FOUND} & A lookup request does not match any payment visible
to the authenticated merchant. \\
\bottomrule
\end{longtable}

\apiblockheading{Happy-path example.}
\begin{lstlisting}[language=json,caption={Create payment request example}]
{
  "order_id": "ORDER-1001",
  "amount": "100000.00",
  "description": "Demo QR payment",
  "ttl_seconds": 900,
  "metadata": {
    "customer_ref": "CUST-1"
  }
}
\end{lstlisting}

\begin{lstlisting}[language=json,caption={Create payment response example}]
{
  "transaction_id": "pay_01J...",
  "order_id": "ORDER-1001",
  "merchant_id": "m_demo",
  "qr_content": "MINI_GATEWAY|merchant_id=m_demo|transaction_id=pay_01J...",
  "qr_image_url": null,
  "qr_image_base64": null,
  "status": "PENDING",
  "expire_at": "2026-04-29T10:15:00Z"
}
\end{lstlisting}
