\subsection{Refunds}

\begin{longtable}{@{}L{0.25\textwidth}L{0.65\textwidth}@{}}
\caption{Refund API family summary}\\
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endfirsthead
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endhead
Purpose & Create full refunds and retrieve refund status for the authenticated
merchant. \\
\bodyrowrule
Intended audience & Merchant backend. \\
\bodyrowrule
Authentication & Merchant HMAC headers. \\
\bodyrowrule
Reference IDs & \apiid{REF-01}, \apiid{REF-02}, \apiid{REF-03}. \\
\bottomrule
\end{longtable}

\apiblockheading{Endpoints.}
\begin{longtable}{@{}L{0.31\textwidth}L{0.59\textwidth}@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endhead
\code{POST /v1/refunds} & Create a refund request against a successful
payment. \\
\bodyrowrule
\code{GET /v1/refunds/{refund_transaction_id}} & Retrieve a refund by gateway
refund transaction identifier. \\
\bodyrowrule
\code{GET /v1/refunds/by-refund-id/{refund_id}} & Retrieve a refund by the
merchant's own refund identifier. \\
\bottomrule
\end{longtable}

\apiblockheading{Important request fields.}
Refund creation requires \code{refund_id}, \code{refund_amount}, and
\code{reason}. The original payment is selected with exactly one of
\code{original_transaction_id} or \code{order_id}.

\apiblockheading{Important response fields.}
The gateway returns \code{refund_transaction_id},
\code{original_transaction_id}, \code{refund_id}, \code{refund_amount}, and
\code{refund_status}.

\apiblockheading{Business rules and usage notes.}
\begin{itemize}
  \item Only full refunds are supported in the current MVP.
  \item The original payment must already be successful before a refund can be
  requested.
  \item The refund amount must exactly match the original payment amount.
  \item Refunds are allowed only within seven days from the payment's success
  timestamp.
  \item Reusing the same merchant refund identifier returns the existing refund
  when it already belongs to the same merchant.
  \item A pending or already successful refund prevents a second active refund
  from being opened for the same completed payment.
  \item A previously failed refund does not block a fresh refund request.
\end{itemize}

\apiblockheading{Likely error codes.}
\begin{longtable}{@{}L{0.30\textwidth}L{0.60\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endhead
\code{MERCHANT_NOT_ACTIVE} & The merchant is not currently allowed to create
refund activity. \\
\bodyrowrule
\code{PAYMENT_NOT_REFUNDABLE} & The referenced payment is not in a state that
can be refunded. \\
\bodyrowrule
\code{REFUND_WINDOW_EXPIRED} & The successful payment is older than the allowed
refund window. \\
\bodyrowrule
\code{REFUND_AMOUNT_NOT_FULL} & The requested refund amount does not exactly
match the original payment amount. \\
\bodyrowrule
\code{REFUND_NOT_ALLOWED} & The refund request violates another refund business
rule, such as a conflicting existing refund. \\
\bodyrowrule
\code{REFUND_NOT_FOUND} & A refund lookup does not match any refund visible to
the authenticated merchant. \\
\bottomrule
\end{longtable}

\apiblockheading{Happy-path example.}
\begin{lstlisting}[language=json,caption={Create refund request example}]
{
  "original_transaction_id": "pay_01J...",
  "refund_id": "REF-1001",
  "refund_amount": "100000.00",
  "reason": "Customer requested refund"
}
\end{lstlisting}

\begin{lstlisting}[language=json,caption={Create refund response example}]
{
  "refund_transaction_id": "rfnd_01J...",
  "original_transaction_id": "pay_01J...",
  "refund_id": "REF-1001",
  "refund_amount": "100000.00",
  "refund_status": "REFUND_PENDING"
}
\end{lstlisting}
