\section{Callback And Webhook Flows}

These flows are asynchronous. The gateway receives result notifications from a
provider-side source and, separately, sends final-state notifications to the
merchant's webhook endpoint.

\subsection{Provider Callbacks}

\begin{longtable}{@{}L{0.25\textwidth}L{0.65\textwidth}@{}}
\caption{Provider callback family summary}\\
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endfirsthead
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endhead
Purpose & Apply payment and refund result updates from a controlled callback
source while preserving evidence for review. \\
\bodyrowrule
Intended audience & Provider callback source in the agreed private or
simulated environment. \\
\bodyrowrule
Authentication & Controlled-environment trust in the current MVP. \\
\bodyrowrule
Reference IDs & \apiid{CB-01}, \apiid{CB-02}, \apiid{REF-04}, \apiid{REF-05}. \\
\bottomrule
\end{longtable}

\apiblockheading{Endpoints.}
\begin{longtable}{@{}L{0.31\textwidth}L{0.59\textwidth}@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endhead
\code{POST /v1/provider/callbacks/payment} & Report the final result of a
payment attempt. \\
\bodyrowrule
\code{POST /v1/provider/callbacks/refund} & Report the final result of a refund
attempt. \\
\bottomrule
\end{longtable}

\apiblockheading{Important request fields.}
Payment callbacks identify a payment with \code{transaction_reference}. Refund
callbacks identify a refund with \code{refund_transaction_id}. Both callback
types carry an external reference, a result status, an amount, a processed
timestamp appropriate to the result type, and an optional raw provider payload
for evidence preservation.

\apiblockheading{Processing results.}
\begin{longtable}{@{}L{0.22\textwidth}L{0.68\textwidth}@{}}
\toprule
\textbf{Result} & \textbf{Meaning} \\
\midrule
\endfirsthead
\toprule
\textbf{Result} & \textbf{Meaning} \\
\midrule
\endhead
\code{PROCESSED} & The callback matched the expected state transition and was
applied. \\
\bodyrowrule
\code{IGNORED} & The callback was a harmless duplicate or otherwise did not
need to change the current state. \\
\bodyrowrule
\code{PENDING_REVIEW} & The callback was accepted as evidence, but the gateway
did not trust it enough to apply the update automatically. \\
\bodyrowrule
\code{FAILED} & The callback could not be processed because the request itself
was not usable. \\
\bottomrule
\end{longtable}

\apiblockheading{Business rules and usage notes.}
\begin{itemize}
  \item Raw callback evidence is preserved before state changes are evaluated.
  \item A pending payment may move to success or failure when a valid callback
  arrives.
  \item A payment that has already expired is not automatically revived by a
  late success callback.
  \item Amount mismatches, unknown references, and conflicting final-state
  callbacks are preserved for review instead of silently overwriting internal
  records.
  \item Refund callbacks follow the same evidence-first pattern and move a
  pending refund to success or failure only when the update is internally
  consistent.
\end{itemize}

\apiblockheading{Likely error and review outcomes.}
\begin{longtable}{@{}L{0.28\textwidth}L{0.62\textwidth}@{}}
\toprule
\textbf{Signal} & \textbf{What it means for the caller} \\
\midrule
\endfirsthead
\toprule
\textbf{Signal} & \textbf{What it means for the caller} \\
\midrule
\endhead
\code{VALIDATION_ERROR} & The callback payload is missing required fields or
uses an invalid shape. \\
\bodyrowrule
\code{PENDING_REVIEW} & The callback was recorded but requires manual review,
usually because the reference, amount, or state did not align with the gateway's
current view. \\
\bottomrule
\end{longtable}

\apiblockheading{Happy-path example.}
\begin{lstlisting}[language=json,caption={Payment callback request example}]
{
  "external_reference": "bank_ref_1001",
  "transaction_reference": "pay_01J...",
  "status": "SUCCESS",
  "amount": "100000.00",
  "paid_at": "2026-04-29T10:00:00Z",
  "source_type": "SIMULATOR",
  "raw_payload": {
    "provider_status": "00"
  }
}
\end{lstlisting}

\begin{lstlisting}[language=json,caption={Payment callback response example}]
{
  "transaction_id": "pay_01J...",
  "status": "SUCCESS",
  "processing_result": "PROCESSED",
  "reconciliation_record_id": null
}
\end{lstlisting}
