\subsection{Outbound Webhooks}

\begin{longtable}{@{}L{0.25\textwidth}L{0.65\textwidth}@{}}
\caption{Outbound webhook family summary}\\
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endfirsthead
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endhead
Purpose & Notify the merchant backend when a payment or refund reaches a final
state. \\
\bodyrowrule
Intended audience & Merchant backend that receives webhook calls from the
gateway. \\
\bodyrowrule
Authentication & Signed delivery headers generated from the merchant's active
credential secret. \\
\bodyrowrule
Reference IDs & \apiid{WH-01} to \apiid{WH-09}. \\
\bottomrule
\end{longtable}

\apiblockheading{Delivery model.}
\begin{longtable}{@{}L{0.27\textwidth}L{0.63\textwidth}@{}}
\toprule
\textbf{Topic} & \textbf{Contract} \\
\midrule
\endfirsthead
\toprule
\textbf{Topic} & \textbf{Contract} \\
\midrule
\endhead
Event types & \code{payment.succeeded}, \code{payment.failed},
\code{payment.expired}, \code{refund.succeeded}, and \code{refund.failed}. \\
\bodyrowrule
Headers & \code{X-Webhook-Event-Id}, \code{X-Webhook-Event-Type},
\code{X-Webhook-Timestamp}, and \code{X-Webhook-Signature}. \\
\bodyrowrule
Success rule & Any HTTP 2xx response from the merchant endpoint marks the
delivery attempt as successful. \\
\bodyrowrule
Automatic retries & Failed delivery attempts are retried after approximately 1,
5, and 15 minutes before the event becomes operationally failed. \\
\bodyrowrule
Business impact & Delivery failure never rolls back a payment or refund final
state. The webhook informs the merchant; it does not decide the transaction
outcome. \\
\bottomrule
\end{longtable}

\apiblockheading{Merchant responsibilities.}
\begin{itemize}
  \item Verify the webhook signature and timestamp before accepting the event.
  \item Treat delivery as at-least-once and make the receiving endpoint
  idempotent.
  \item Use the event identifier to detect duplicates safely.
  \item Fall back to payment or refund query APIs if local processing is
  uncertain.
\end{itemize}

\apiblockheading{Happy-path example.}
\begin{lstlisting}[language=json,caption={Webhook event example}]
{
  "event_id": "evt_01J...",
  "event_type": "payment.succeeded",
  "merchant_id": "m_demo",
  "entity_type": "PAYMENT",
  "entity_id": "payment-row-uuid",
  "created_at": "2026-04-29T10:05:01Z",
  "data": {
    "transaction_id": "pay_01J...",
    "order_id": "ORDER-1001",
    "status": "SUCCESS",
    "amount": "100000.00"
  }
}
\end{lstlisting}

If delivery retries are exhausted, the event remains available for internal
support recovery. The manual retry capability is documented later as a
restricted operational action rather than a merchant-facing API.
