\section{Merchant Dashboard APIs}

The Merchant Dashboard is an interactive read-oriented surface for human users.
It is separate from the merchant backend API. Dashboard users sign in with a
session account, not with HMAC credentials.

\begin{longtable}{@{}L{0.25\textwidth}L{0.65\textwidth}@{}}
\caption{Merchant Dashboard API family summary}\\
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endfirsthead
\toprule
\textbf{Item} & \textbf{Details} \\
\midrule
\endhead
Purpose & Let merchant staff review payments, refunds, webhooks, profile data,
and analytics through a signed-in dashboard session. \\
\bodyrowrule
Intended audience & Merchant dashboard users. \\
\bodyrowrule
Authentication & Merchant dashboard session cookie. \\
\bodyrowrule
Reference IDs & \apiid{AUTH-02}, \apiid{AUTH-03}, \apiid{MP-01} to \apiid{MP-06}. \\
\bottomrule
\end{longtable}

\subsection{Session And Overview}

\begin{longtable}{@{}L{0.33\textwidth}L{0.57\textwidth}@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endhead
\routecell{\code{POST /v1/merchant-portal/auth/login}\\
\code{POST /v1/merchant-portal/auth/logout}\\
\code{GET /v1/merchant-portal/auth/me}\\
\code{POST /v1/merchant-portal/auth/change-password}} & Manage dashboard
session lifecycle for a merchant user. \\
\bodyrowrule
\routecell{\code{GET /v1/merchant-portal/dashboard/summary}\\
\code{GET /v1/merchant-portal/dashboard/charts}} & Return merchant-scoped
overview counters and chart series for the signed-in user. \\
\bottomrule
\end{longtable}

\subsection{Analytics, Explorers, And Profile}

\begin{longtable}{@{}L{0.33\textwidth}L{0.57\textwidth}@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endfirsthead
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\endhead
\code{GET /v1/merchant-portal/analytics?days=7|30|90} & Return merchant-scoped
analytics for a selected reporting window. \\
\bodyrowrule
\routecell{\code{GET /v1/merchant-portal/payments}\\
\code{GET /v1/merchant-portal/payments/{transaction_id}}} & List payments and
open payment detail for the signed-in merchant. \\
\bodyrowrule
\routecell{\code{GET /v1/merchant-portal/refunds}\\
\code{GET /v1/merchant-portal/refunds/{refund_transaction_id}}} & List refunds
and open refund detail for the signed-in merchant. \\
\bodyrowrule
\routecell{\code{GET /v1/merchant-portal/webhooks}\\
\code{GET /v1/merchant-portal/webhooks/{event_id}}} & List webhook events and
open webhook detail for the signed-in merchant. \\
\bodyrowrule
\routecell{\code{GET /v1/merchant-portal/profile}\\
\code{GET /v1/merchant-portal/credentials}} & Return read-only merchant profile
metadata and credential metadata. \\
\bottomrule
\end{longtable}

\apiblockheading{Business rules and usage notes.}
\begin{itemize}
  \item Dashboard session scope is tied to the signed-in user and their merchant
  account. Clients do not send an extra merchant identifier to choose data
  scope after login.
  \item Analytics accepts only three reporting windows: 7, 30, and 90 days.
  \item Zero-activity dates still appear in analytics responses so charts remain
  visually stable across the selected range.
  \item Detail routes return not found when a record does not belong to the
  signed-in merchant, even if the identifier exists elsewhere in the system.
  \item Credential metadata may be displayed, but raw credential secrets are
  never returned by the dashboard APIs.
\end{itemize}

\apiblockheading{Likely error codes.}
\begin{longtable}{@{}L{0.31\textwidth}L{0.59\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{When it matters} \\
\midrule
\endhead
\code{MERCHANT_AUTH_INVALID_CREDENTIALS} & The sign-in request uses an unknown
merchant identifier, email, or password combination. \\
\bodyrowrule
\code{MERCHANT_AUTH_INACTIVE} & The dashboard account exists but is not
currently active. \\
\bodyrowrule
\code{MERCHANT_AUTH_REQUIRED} or \code{MERCHANT_AUTH_INVALID_SESSION} & The
request does not include a usable signed-in dashboard session. \\
\bodyrowrule
\code{INVALID_ANALYTICS_RANGE} & The analytics request asks for a reporting
window other than 7, 30, or 90 days. \\
\bodyrowrule
\code{PAYMENT_NOT_FOUND}, \code{REFUND_NOT_FOUND}, \code{WEBHOOK_EVENT_NOT_FOUND}
& A detail request refers to a record outside the user's merchant scope or to a
record that does not exist. \\
\bottomrule
\end{longtable}

\apiblockheading{Happy-path example.}
\begin{lstlisting}[language=json,caption={Merchant dashboard login request example}]
{
  "merchant_id": "m_demo",
  "email": "merchant@example.com",
  "password": "temporary-or-user-password"
}
\end{lstlisting}

\begin{lstlisting}[language=json,caption={Merchant dashboard login response example}]
{
  "user": {
    "user_id": "uuid",
    "merchant_id": "m_demo",
    "email": "merchant@example.com",
    "full_name": "Merchant Admin",
    "role": "MERCHANT_ADMIN",
    "status": "ACTIVE"
  },
  "merchant_status": "ACTIVE"
}
\end{lstlisting}
