\section{Error Catalogue}

All API error responses use the same envelope:

\begin{lstlisting}[language=json,caption={Shared error response shape}]
{
  "error_code": "PAYMENT_ALREADY_SUCCESS",
  "message": "Payment already succeeded for this order.",
  "request_id": "req_01J...",
  "details": {}
}
\end{lstlisting}

\docbox{How to read the tables below}{%
\begin{itemize}
  \item \textbf{Meaning and common trigger} explains both what the code means
  and the situation in which it usually appears.
  \item \textbf{Retry?} answers whether sending the same request again is
  normally useful.
  \item \textbf{What to do next} points to the next action for the caller.
  \item Callback endpoints sometimes return a normal HTTP response with
  \code{processing_result = PENDING_REVIEW} instead of raising a hard error. In
  those cases the callback was recorded as evidence but not applied
  automatically.
\end{itemize}
}

\subsection{Merchant Backend Authentication And Validation}

{\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}L{0.21\textwidth}L{0.07\textwidth}L{0.36\textwidth}L{0.08\textwidth}L{0.18\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endhead
\code{VALIDATION_ERROR} & 422 & The request shape or a field value is invalid,
typically because a required field is missing or an unsupported value was sent.
& No & Correct the request and submit again. \\
\bodyrowrule
\code{AUTH_MISSING_HEADER} & 401 & A required HMAC header is missing from the
merchant request. & No & Add the missing header and sign a fresh request. \\
\bodyrowrule
\code{AUTH_INVALID_MERCHANT} & 401 & The merchant identifier is unknown to the
gateway. & No & Confirm the merchant id supplied during onboarding. \\
\bodyrowrule
\code{AUTH_INVALID_CREDENTIAL} & 401 & The access key is wrong, inactive, or
belongs to another merchant. & No & Use the currently active credential set. \\
\bodyrowrule
\code{AUTH_TIMESTAMP_EXPIRED} & 401 & The signed timestamp is outside the
allowed window, usually because of clock skew or replay of an old request. &
Usually no & Synchronize the server clock and sign a new request. \\
\bodyrowrule
\code{AUTH_INVALID_SIGNATURE} & 401 & The signature does not match the gateway's
calculation because the signed method, path, body, or secret differs from the
actual request. & No & Rebuild the canonical signing string and sign again. \\
\bottomrule
\end{longtable}
}

\subsection{Merchant Payment And Refund Business Rules}

{\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}L{0.21\textwidth}L{0.07\textwidth}L{0.36\textwidth}L{0.08\textwidth}L{0.18\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endhead
\code{MERCHANT_NOT_ACTIVE} & 403 & The merchant is not currently allowed to
transact because the account is still under review, suspended, or disabled. &
No & Contact onboarding or support. \\
\bodyrowrule
\code{PAYMENT_PENDING_EXISTS} & 409 & A conflicting pending payment already
exists for the order. & Usually no & Reuse or inspect the existing pending
payment. \\
\bodyrowrule
\code{PAYMENT_ALREADY_SUCCESS} & 409 & The order has already been paid
successfully and should not be reopened as a new payment. & No & Treat the
order as paid and query the existing payment if needed. \\
\bodyrowrule
\code{PAYMENT_NOT_FOUND} & 404 & The payment lookup identifier is wrong or does
not belong to the caller's merchant scope. & No & Recheck the payment
identifier before retrying. \\
\bodyrowrule
\code{PAYMENT_NOT_REFUNDABLE} & 409 & The referenced payment is not in a state
that can be refunded. & No & Query the payment state and confirm refund
eligibility. \\
\bodyrowrule
\code{REFUND_WINDOW_EXPIRED} & 409 & The successful payment is older than the
allowed seven-day refund window. & No & Resolve the case through support. \\
\bodyrowrule
\code{REFUND_AMOUNT_NOT_FULL} & 409 & The refund amount does not match the full
original payment amount, which means a partial refund was attempted. & No &
Submit a full refund or wait for a future phase that supports partial refunds. \\
\bodyrowrule
\code{REFUND_NOT_ALLOWED} & 409 & Another refund rule blocks the request, such
as an incompatible existing refund flow. & Usually no & Query current refund
state before trying another refund. \\
\bodyrowrule
\code{REFUND_NOT_FOUND} & 404 & The refund identifier is wrong or outside the
caller's merchant scope. & No & Recheck the refund identifier before retrying. \\
\bottomrule
\end{longtable}
}

\subsection{Merchant Dashboard Session And Scope Errors}

{\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}L{0.21\textwidth}L{0.07\textwidth}L{0.36\textwidth}L{0.08\textwidth}L{0.18\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endhead
\code{MERCHANT_AUTH_REQUIRED} & 401 & The dashboard request needs a signed-in
session, but no session cookie was supplied. & Yes & Sign in and repeat the
action. \\
\bodyrowrule
\code{MERCHANT_AUTH_INVALID_SESSION} & 401 & The dashboard session expired,
was replaced, or is otherwise invalid. & Yes & Sign in again and retry. \\
\bodyrowrule
\code{MERCHANT_AUTH_INVALID_CREDENTIALS} & 401 & The merchant id, email, or
password does not match an active dashboard user. & No & Recheck credentials or
request a password reset. \\
\bodyrowrule
\code{MERCHANT_AUTH_INACTIVE} & 403 & The dashboard user exists but is not
active. & No & Ask support to review the dashboard account status. \\
\bodyrowrule
\code{INVALID_ANALYTICS_RANGE} & 422 & The analytics request used a reporting
window other than 7, 30, or 90 days. & No & Use one of the supported windows. \\
\bottomrule
\end{longtable}
}

\subsection{Restricted Internal And Operational Errors}

{\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}L{0.21\textwidth}L{0.07\textwidth}L{0.36\textwidth}L{0.08\textwidth}L{0.18\textwidth}@{}}
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endfirsthead
\toprule
\textbf{Code} & \textbf{HTTP} & \textbf{Meaning and common trigger} &
\textbf{Retry?} & \textbf{What to do next} \\
\midrule
\endhead
\code{INTERNAL_AUTH_REQUIRED} & 401 & A restricted endpoint was called without
an internal operator session. & Yes & Sign in as an internal operator. \\
\bodyrowrule
\code{INTERNAL_AUTH_INVALID_SESSION} & 401 & The internal session expired, was
replaced, or is malformed. & Yes & Sign in again before retrying. \\
\bodyrowrule
\code{INTERNAL_AUTH_INVALID_CREDENTIALS} & 401 & The internal operator sign-in
credentials are incorrect. & No & Correct the credentials or reset the
password. \\
\bodyrowrule
\code{INTERNAL_AUTH_INACTIVE} & 403 & The internal operator account exists but
is disabled. & No & Ask an administrator to reactivate the account. \\
\bodyrowrule
\code{INTERNAL_AUTH_FORBIDDEN} & 403 & The signed-in operator lacks the role
required for the action. & No & Use an account with the right role or hand the
task to the proper operator. \\
\bodyrowrule
\code{INTERNAL_AUTH_BOOTSTRAP_DISABLED} & 409 & First-time bootstrap is already
closed because an administrator exists. & No & Use normal internal sign-in. \\
\bodyrowrule
\code{INTERNAL_USER_ALREADY_EXISTS} & 409 & The internal user being created
already exists. & No & Update the existing user instead. \\
\bodyrowrule
\code{INTERNAL_USER_NOT_FOUND} & 404 & The targeted internal user record does
not exist. & No & Recheck the internal user identifier. \\
\bodyrowrule
\code{LAST_ACTIVE_ADMIN_REQUIRED} & 409 & The requested change would remove the
final active administrator. & No & Keep at least one active administrator. \\
\bodyrowrule
\code{MERCHANT_ALREADY_EXISTS} & 409 & Merchant registration reused an existing
merchant identifier. & No & Choose the correct existing merchant or a new
identifier. \\
\bodyrowrule
\code{MERCHANT_NOT_FOUND} & 404 & The targeted merchant record does not exist.
& No & Recheck the merchant identifier. \\
\bodyrowrule
\code{ONBOARDING_CASE_NOT_FOUND} & 404 & Approval, rejection, or detail read was
attempted before an onboarding case existed. & No & Submit an onboarding case
first. \\
\bodyrowrule
\code{ONBOARDING_CASE_FINAL} & 409 & The onboarding case is already approved or
rejected and cannot be overwritten. & No & Review the existing decision
instead. \\
\bodyrowrule
\code{ONBOARDING_CASE_NOT_APPROVED} & 409 & Merchant activation was requested
before onboarding approval. & No & Complete onboarding review first. \\
\bodyrowrule
\code{ACTIVE_CREDENTIAL_EXISTS} & 409 & A merchant already has an active
credential, so first-time credential creation is blocked. & No & Rotate or
inspect the current credential instead. \\
\bodyrowrule
\code{ACTIVE_CREDENTIAL_NOT_FOUND} & 404 & Credential rotation was requested
before any active credential existed. & No & Create the first active credential
first. \\
\bodyrowrule
\code{ACTIVE_CREDENTIAL_REQUIRED} & 409 & Merchant activation needs an active
credential, but none exists. & No & Create a credential before activating the
merchant. \\
\bodyrowrule
\code{MERCHANT_PORTAL_USER_ALREADY_EXISTS} & 409 & The dashboard user already
exists for the merchant. & No & Update the existing dashboard user instead. \\
\bodyrowrule
\code{MERCHANT_PORTAL_USER_NOT_FOUND} & 404 & The targeted dashboard user does
not exist. & No & Recheck the dashboard user identifier. \\
\bodyrowrule
\code{RECONCILIATION_NOT_FOUND} & 404 & The reconciliation record identifier is
unknown. & No & Recheck the reconciliation record identifier. \\
\bodyrowrule
\code{RECONCILIATION_ALREADY_RESOLVED} & 409 & The mismatch record is already
closed. & No & Review the resolved record instead of resending the action. \\
\bodyrowrule
\code{WEBHOOK_EVENT_NOT_FOUND} & 404 & The webhook event record does not exist.
& No & Recheck the webhook event identifier. \\
\bodyrowrule
\code{WEBHOOK_RETRY_NOT_ALLOWED} & 409 & The webhook event is not currently in a
failed state and cannot be retried manually. & No & Inspect the webhook event
status before retrying. \\
\bottomrule
\end{longtable}
}
