\section{Security And Go-Live Notes}

\subsection{Security Controls Already Enforced}

\begin{itemize}
  \item Merchant backend payment and refund APIs require HMAC authentication.
  \item Merchant Dashboard sessions and restricted internal sessions are
  separated from each other and from HMAC credentials.
  \item Merchant Dashboard data scope is derived from the signed-in dashboard
  account instead of trusting a client-supplied merchant selector.
  \item Raw merchant credential secrets are not returned by read APIs after
  issuance.
  \item Generated passwords for dashboard users and internal users are returned
  only during create or reset actions.
  \item Operational write actions preserve a human reason for later audit.
  \item Callback evidence is preserved even when the gateway cannot safely apply
  the requested update automatically.
\end{itemize}

\subsection{Go-Live Expectations}

\begin{itemize}
  \item Protect the merchant shared secret and access key as production-grade
  credentials.
  \item Verify outbound webhook signatures and timestamps before accepting
  events.
  \item Make webhook receivers idempotent because delivery is at-least-once.
  \item Keep the provider callback source in a controlled private environment
  unless additional internet-facing hardening is added later.
  \item Establish an operational process for webhook recovery and reconciliation
  review so ambiguous cases are handled consistently.
  \item Treat the Merchant Dashboard as a human review surface rather than as a
  substitute for server-to-server payment integration.
\end{itemize}

\docbox{Merchant go-live checklist}{%
\begin{enumerate}[leftmargin=1.7em]
  \item Confirm the merchant identifier, access key, and shared secret were
  received securely.
  \item Verify the merchant backend can generate valid HMAC signatures for
  payment and refund requests.
  \item Confirm the merchant webhook endpoint returns HTTP 2xx only after the
  event is durably accepted.
  \item Test duplicate webhook handling and payment-query fallback logic.
  \item Confirm the Merchant Dashboard account can sign in and that the portal
  password has been rotated away from any temporary onboarding value.
  \item Align with internal support on who handles failed webhook recovery and
  reconciliation review.
\end{enumerate}
}

The current MVP deliberately keeps provider callback trust inside controlled
environments and keeps several advanced operational hardening steps outside the
documented scope. Teams preparing for a broader production launch should plan a
separate hardening review before exposing these flows to open internet traffic.
